---
title: "<center>Behavioral Data Science<br>Project - Part I</center>"
author: "<center>Avisek Choudhury<br>Ben Corriveau<br>Cindy Nikolai</center>"
date: "<center>10/12/2020</center>"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Research Questions 

1.  Is there any correlation between overall feeling of safety and age and/or income and/or gender and/or frequency of travel?  
2.  Is there any pattern between the satisfaction with various SFO facilities/services and income and/or gender and/or age?  
3.  Does there appear to be any pattern between the purpose of your trip today and the overall feeling of cleanliness at the airport?  

We considered that comments are correlated with ratings for some of the questions.  We will wait for further analysis before determining this.


```{r readData, message=FALSE, warning=FALSE, error=TRUE}
#Import the libraries
library(tidyverse)
library(haven)
library(plotly)
library(reshape2)

#Read the data set
sfoDf <- read_delim('SFO_survey_withText.txt', delim="\t")
head(sfoDf)
```

```{r countingNA, message=FALSE, warning=FALSE, error=TRUE }
#Count the NA or missing Rows for each columns
sapply(sfoDf, function(x) sum(is.na(x)))
```

We see lot of columns with huge no of missing values. Let's check the columns with more than 50% values are missing.

```{r countingNA50, message=FALSE, warning=FALSE, error=TRUE }
#Count the NA or missing Rows for each columns
unlist(sapply(sfoDf, function(x) { 
  if (sum(is.na(x)) > length(x)* 0.5)  
    return (sum(is.na(x))) 
  }))
```

So we see there are 40 columns out of 101 columns has more than 50% missing values.

```{r flightByTerminal, message=FALSE, warning=FALSE, error=TRUE}
#Airline terminal
# 1 Terminal 1
# 2 International Terminal
# 3 Terminal 3
# 0 Unknown
flightByTerm <- sfoDf %>% 
  group_by(TERM) %>%
  summarise(Flight_By_Terminal = n()) %>% 
  rename(Terminal = TERM)

flightByTerm
```
```{r flightByTerminalPlot, message=FALSE, warning=FALSE, error=TRUE}
fig <- plot_ly(
  flightByTerm,
  x = ~as.character(Terminal),
  y = ~Flight_By_Terminal,
  type = "bar"
)
fig <- fig %>% 
  layout(title = " ",
         xaxis = list(title = "Terminal No"),
         yaxis = list(title = "No of Flights By Terminal"))

fig
```



```{r flightByCarrier, message=FALSE, warning=FALSE, error=TRUE}
#Airline type (based on Sampling Plan)
# 1 Major carriers
# 2 Small/International carriers
# 3 New carriers
sfoDf %>% 
  group_by(ATYPE) %>%
  summarise(Flight_By_Carrier = n()) %>% 
  rename(Airline_Type = ATYPE)
```

```{r flightByDest, message=FALSE, warning=FALSE, error=TRUE}
# DEST Destination of flight
# 1 Within California
# 2 Out of state
# 3 Out of country
sfoDf %>% 
  group_by(DEST) %>%
  summarise(Flight_By_Destination = n()) %>% 
  rename(Destination = DEST)
```

Lets see the no of flights by each airline type and destination for each terminal.

```{r , message=FALSE, warning=FALSE, error=TRUE}
flightByTCD <- sfoDf %>% 
  rename(Terminal = TERM, Airline_Type = ATYPE, Destination = DEST) %>% 
  group_by(Terminal, Airline_Type, Destination) %>%
  summarise(Flight_Count = n()) 

flightByTCD$Terminal <-  as.factor(flightByTCD$Terminal)
levels(flightByTCD$Terminal) <- c("Terminal 1", "Terminal 2", "Terminal 3")
flightByTCD$Airline_Type <- as.factor(flightByTCD$Airline_Type)
levels(flightByTCD$Airline_Type) <- c("Major carriers", "Small/International carriers", "New carriers")
flightByTCD$Destination <- as.factor(flightByTCD$Destination)

p <- ggplot(flightByTCD, aes(Airline_Type, Flight_Count)) +   
    geom_bar(aes(fill = Destination), 
             position = "dodge", stat = "identity") + 
    labs(title = "Flight Count by Airline Type and Destination for Each Terminal ", 
         x = "\nAirline Type", y = "Flight Count\n", color = "Destination\n") + 
  scale_fill_discrete( name="Destinations",
                       breaks=c("1", "2", "3"),
                       labels=c("Within California","Out of State","Out of Country")) + 
  theme(axis.text.x = element_text(angle = 30, vjust = 0.2, hjust= 0.2))
 
p <- p + facet_wrap( ~ Terminal, ncol=3)

#ggplotly(p)
p
```


Now let's see top 5 busiest gates by terminal.

```{r , message=FALSE, warning=FALSE, error=TRUE}
# sfoDf %>% 
#   group_by(TERM, GATENUM) %>% 
#   summarise(Flight_Count = n()) %>% 
#   arrange(desc(Flight_Count)) %>% 
#   head(n=20) %>% 
#   rename(Terminal = TERM, Gate_Num = GATENUM)

top5GtByTerm <- sfoDf %>% 
  group_by(TERM, GATENUM) %>% 
  summarise(Flight_Count = n()) %>% 
  arrange(desc(Flight_Count)) %>% 
  top_n(n=5) %>% 
  arrange(TERM, GATENUM) %>% 
  rename(Terminal = TERM, Gate_Num = GATENUM)

top5GtByTerm

p <- ggplot(top5GtByTerm, aes(Gate_Num, Flight_Count)) + 
  geom_bar(position = "dodge", stat = "identity") + 
  geom_text(aes(label=Flight_Count), vjust=0) + 
  facet_wrap( ~ Terminal, ncol=3)

p

```


Let's take a look at cleanliness responses grouped by various attributes (gender, income, age, etc.)
```{r}
cleanliness_data <- sfoDf %>%
  select('Q17','Q18':'Q19','Q8COM1':'Q8COM2') %>% 
  rename(Age = Q17,
         Gender = Q18,
         Income = Q19) %>% 
  melt(., id.vars=c('Age','Gender','Income'), value.name = 'Comments') %>% 
  select(-'variable') %>% 
  na.omit(.)

# Change numeric to real answers - Age
cleanliness_data$Age[cleanliness_data$Age==1] <- 'Under 18'
cleanliness_data$Age[cleanliness_data$Age==2] <- '18 - 24'
cleanliness_data$Age[cleanliness_data$Age==3] <- '25 - 34'
cleanliness_data$Age[cleanliness_data$Age==4] <- '35 - 44'
cleanliness_data$Age[cleanliness_data$Age==5] <- '45 - 54'
cleanliness_data$Age[cleanliness_data$Age==6] <- '55 - 64'
cleanliness_data$Age[cleanliness_data$Age==7] <- '65 and over'
cleanliness_data$Age[cleanliness_data$Age==8] <- "Don't know/Refused"
cleanliness_data$Age[cleanliness_data$Age==0] <- 'Blank'

# Change numeric to real answers - Gender
cleanliness_data$Gender[cleanliness_data$Gender==1] <- 'Male'
cleanliness_data$Gender[cleanliness_data$Gender==2] <- 'Female'
cleanliness_data$Gender[cleanliness_data$Gender==0] <- 'Blank'

# Change numeric to real answers - Income
cleanliness_data$Income[cleanliness_data$Income==1] <- 'Under $50,000'
cleanliness_data$Income[cleanliness_data$Income==2] <- '$50,000-$100,000'
cleanliness_data$Income[cleanliness_data$Income==3] <- '$100,001-$150,000'
cleanliness_data$Income[cleanliness_data$Income==4] <- 'Over $150,000'
cleanliness_data$Income[cleanliness_data$Income==5] <- 'Other (unspecified)'
cleanliness_data$Income[cleanliness_data$Income==0] <- 'Blank'
```

```{r}
ggplot(cleanliness_data, aes(x=as.factor(Comments), fill=Income)) + 
  geom_bar(stat='count') +
  facet_wrap(~Gender) + 
  labs(title="Cleanliness Comments by Gender/Income")
```

We can see that the most common responses within the cleanliness comments were 3 - "Restroom dirty/not cleaned often enough/wet floor," and 12 - "General positive comment."  Comments 4, 5, and 6 were also common responses, which are all related to bathroom cleanliness.  The largest number of respondendts appears to have income within $50,000 - $100,00. This could be because they are more critical, or because they are the largest respondent group.

```{r}
ggplot(cleanliness_data, aes(x=as.factor(Comments), fill=Age)) + 
  geom_bar(stat='count') +
  facet_wrap(~Gender) +
  labs(title="Cleanliness Comments by Gender/Age")
```
We can see here that the group with the most critical responses for bathroom cleanliness appears to be aged 25-34.  This could be because they are more critical, or because they are the largest respondent group.

Let's take a look at the counts of responses by Age, Gender, and Income.
```{r}
ggplot(cleanliness_data, aes(x=Age)) + 
  geom_bar(stat='count') +
  facet_wrap(~Gender) +
  labs(title="Cleanliness Comment Count by Age and Gender")
```


```{r}
ggplot(cleanliness_data, aes(x=Income)) + 
  geom_bar(stat='count') +
  facet_wrap(~Gender) +
  labs(title="Cleanliness Comment Count by Income and Gender")+ 
  theme(axis.text.x = element_text(angle = 30, vjust = 0.7))
```
